---
title: "String Manipulation in R with stringr"
author: "QuantiDot"
date: "December 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# String basics

Let's start with some basics: how to enter strings in R, how to control how numbers are transformed to strings, and finally how to combine strings together to produce output that combines text and nicely formatted numbers.

## Double quotes vs single quotes
Prefer double quotes (") to single quotes ('). 
There are cases where you need both `'` and `"` inside the string. In this case, fall back to the first guideline and use " to define the string, but you'll have to escape any double quotes inside the string using a backslash (i.e. `\"`).

```{r}
# Define line1
line1 <- "The table was a large one, but the three were all crowded together at one corner of it:"
# Define line2
line2 <- "\"No room! No room!\" they cried out when they saw Alice coming."
# Define line3
line3 <- "\"There's plenty of room!\" said Alice indignantly, and she sat down in a large arm-chair at one end of the table."

```
```{r, echo = FALSE}
line1
line2
line3

```

To avoid printing `"\"` then use `writeLines()`
```{r}
# Putting lines in a vector
lines <- c(line1, line2, line3)

# Print lines
lines

# Use writeLines() on lines
writeLines(lines)

# Write lines with a space separator
writeLines(lines, sep = "")
```
```{r}
# Should display: To have a \ you need \\
writeLines('To have a \\ you need \\\\')

# Should display: 
# This is a really 
# really really 
# long string
writeLines("This is a really\n\ really really\n\ long string")
```

## Turning numbers into strings
```{r}
# Some vectors of numbers
percent_change  <- c(4, -1.91, 3.00, -5.002)
income <-  c(72.19, 1030.18, 10291.93, 1189192.18)
p_values <- c(0.12, 0.98, 0.0000191, 0.00000000002)

# Format c(0.0011, 0.011, 1) with digits = 1
format(c(0.0011, 0.011, 1),  digits = 1)

# Format c(1.0011, 2.011, 1) with digits = 1
format(c(1.0011, 2.011, 1),  digits = 1)

# Format percent_change to one place after the decimal point
format(percent_change, digit = 2)

# Format income to whole numbers
format(income, digits = 2)

# Format p_values in fixed format
format(p_values, scientific = F)
```

Not only does format() control the way the number is represented, it also controls some of the properties of the resulting string that affect its display.

```{r}
# Some vectors of numbers
percent_change  <- c(4, -1.91, 3.00, -5.002)
income <-  c(72.19, 1030.18, 10291.93, 1189192.18)
p_values <- c(0.12, 0.98, 0.0000191, 0.00000000002)

# Format c(0.0011, 0.011, 1) with digits = 1
format(c(0.0011, 0.011, 1),  digits = 1)

# Format c(1.0011, 2.011, 1) with digits = 1
format(c(1.0011, 2.011, 1),  digits = 1)

# Format percent_change to one place after the decimal point
format(percent_change, digit = 2)

# Format income to whole numbers
format(income, digits = 2)
```

The function formatC() provides an alternative way to format numbers based on C style syntax.

Rather than a scientific argument, formatC() has a format argument that takes a code representing the required format. The most useful are:

"f" for fixed,
"e" for scientific, and
"g" for fixed unless scientific saves space

```{r}
# From the format() exercise
x <- c(0.0011, 0.011, 1)
y <- c(1.0011, 2.011, 1)

# formatC() on x with format = "f", digits = 1
formatC(x, format = "f", digits = 1)

# formatC() on y with format = "f", digits = 1
formatC(y, format = "f", digits = 1)

# Format percent_change to one place after the decimal point
formatC(percent_change, format = "f", digits = 1)

# percent_change with flag = "+"
#The flag argument allows you to provide some modifiers that, for example, force
#the display of the sign (flag = "+"), left align numbers (flag = "-") and pad
#numbers with leading zeros (flag = "0"). You'll see an example in this exercise.

formatC(percent_change, format = "f", digits = 1, flag = "+")

# Format p_values using format = "g" and digits = 2
formatC(p_values, format = "g", digits = 2)
```

# Putting strings together with paste()
```{r}
pretty_income <- c("72.19", "1030.18", "10291.93", "1189192.18")
pretty_percent <- c("+4.0", "-1.9", "+3.0", "-5.0")
# Add $ to pretty_income
paste("$", pretty_income, sep = "")
 
# Add % to pretty_percent
paste(pretty_percent, "%", sep = "")

# Create vector with elements like 2010: +4.0%`
years <- c(2010, 2011, 2012, 2013)
year_percent <- paste(years, sep = ": ", paste(pretty_percent,"%", sep = ""))

# Collapse all years into single string
paste(year_percent, collapse = ",")
```

# A very simple table
```{r}
# Define the names vector
income <- c(72.19, 1030.18, 10291.93, 1189192.18)
income_names <- c("Year 0", "Year 1", "Year 2", "Project Lifetime")


# Create pretty_income by using format() with digits = 2 and big.mark = "," (income is loaded)
pretty_income <- format(income, digits = 2 , big.mark = ",")

# Create dollar_income
dollar_income <- paste("$", pretty_income, sep = "")

# Create formatted_names
formatted_names <- format(income_names, justify = "right")

# Create rows
rows <- paste(formatted_names, dollar_income, sep = "   ")

# Write rows
writeLines(rows)
```

# another example
```{r}
toppings <- c("anchovies", "artichoke", "bacon", "breakfast bacon",
 "Canadian bacon", "cheese", "chicken", "chili peppers",
 "feta", "garlic", "green peppers", "grilled onions",  
"ground beef", "ham", "hot sauce", "meatballs", "mushrooms", "olives", "pineapple", "sausage", "spinach", "sun-dried tomato", "tomatoes")

# Randomly sample 3 toppings
my_toppings <- sample(toppings, size = 3)

# Print my_toppings
my_toppings

# Paste "and " to last element: my_toppings_and
my_toppings_and <- paste(c("", "", "and "), my_toppings, sep = "")

# Collapse with comma space: these_toppings
these_toppings <- paste(my_toppings_and, collapse = ", ")

# Add rest of sentence: my_order
my_order <- paste("I want to order a pizza with ", these_toppings, ".", sep = "")

# Order pizza with writeLines()
writeLines(my_order)
```



# Introducing stringr
```{r}
library(stringr)
#Let's look at str_c(), the c is short for concatenate, a function that works like paste()
my_toppings <- c("cheese", NA, NA)
my_toppings_and <- paste(c("", "", "and "), my_toppings, sep = "")
my_toppings_and

# Use str_c() instead of paste(): my_toppings_str
my_toppings_str <- str_c(c("", "", "and "), my_toppings)

# Print my_toppings_str
my_toppings_str
# Another stringr function that is useful when you may have missing values, is str_replace_na() which replaces missing values with any string you choose.

# paste() my_toppings_and with collapse = ", "
paste(my_toppings_and, collapse = ", ")
```

__str_length()__ takes a vector of strings as input and returns the number of characters in each string. 
```{r}

library(stringr)
library(babynames)
library(dplyr)

# Extracting vectors for boys' and girls' names
babynames_2014 <- filter(babynames, year == 2014)
boy_names <- filter(babynames_2014, sex == "M")$name
girl_names <- filter(babynames_2014, sex == "F")$name

# Take a look at a few boy_names
head(boy_names)

# Find the length of all boy_names
boy_length <- str_length(boy_names)

# Take a look at a few lengths
head(boy_length)

# Find the length of all girl_names
girl_length <- str_length(girl_names)

# Find the difference in mean length
mean(girl_length) - mean(boy_length)

# Confirm str_length() works with factors
head(str_length(factor(boy_names)))
```
```{r}
#str_sub() function in stringr extracts parts of strings based on their location.
# Extract first letter from boy_names
boy_first_letter <- str_sub(boy_names, 1, 1)

# Tabulate occurrences of boy_first_letter
table(boy_first_letter)
  
# Extract the last letter in boy_names, then tabulate
boy_last_letter <- str_sub(boy_names, -1, -1)
table(boy_last_letter)

# Extract the first letter in girl_names, then tabulate
girl_first_letter <- str_sub(girl_names, 1, 1)
table(girl_first_letter)

# Extract the last letter in girl_names, then tabulate
girl_last_letter <- str_sub(girl_names, -1, -1)
table(girl_last_letter)
```


__str_detect()__ is used to answer the question: Does the string contain the pattern? It returns a logical vector of the same length as that of the input vector string

```{r}
# Look for pattern "zz" in boy_names
contains_zz <- str_detect(boy_names, fixed("zz"))

# Examine str() of contains_zz
str(contains_zz)

# How many names contain "zz"?
sum(contains_zz)

# Which names contain "zz"?
boy_names[contains_zz]
```

__str_subset__  gets a new vector of strings, but it only contains those original strings that contained the pattern

```{r}
#Find boy_names that contain "zz"
str_subset(boy_names, "zz")

# Find girl_names that contain "zz"
str_subset(girl_names, "zz")

# Find girl_names that contain "U"
starts_U <- str_subset(girl_names, "U")
starts_U

# Find girl_names that contain "U" and "z"
str_subset(starts_U, "z")
```

__str_count()__` answers the question "How many times does the pattern occur in each string?".

# Pattern matching with regular expressions
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
