---
title: "String Manipulation in R with stringr"
author: "QuantiDot"
date: "December 29, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# String basics

Let's start with some basics: how to enter strings in R, how to control how numbers are transformed to strings, and finally how to combine strings together to produce output that combines text and nicely formatted numbers.

###### Double quotes vs single quotes
Prefer double quotes (") to single quotes ('). 
There are cases where you need both `'` and `"` inside the string. In this case, fall back to the first guideline and use " to define the string, but you'll have to escape any double quotes inside the string using a backslash (i.e. `\"`).

```{r}
# Define line1
line1 <- "The table was a large one, but the three were all crowded together at one corner of it:"
# Define line2
line2 <- "\"No room! No room!\" they cried out when they saw Alice coming."
# Define line3
line3 <- "\"There's plenty of room!\" said Alice indignantly, and she sat down in a large arm-chair at one end of the table."

```
```{r, echo = FALSE}
line1
line2
line3

```

To avoid printing `"\"` then use `writeLines()`
```{r}
# Putting lines in a vector
lines <- c(line1, line2, line3)

# Print lines
lines

# Use writeLines() on lines
writeLines(lines)

# Write lines with a space separator
writeLines(lines, sep = "")
```
```{r}
# Should display: To have a \ you need \\
writeLines('To have a \\ you need \\\\')

# Should display: 
# This is a really 
# really really 
# long string
writeLines("This is a really\n\ really really\n\ long string")
```

###### Turning numbers into strings
```{r}
# Some vectors of numbers
percent_change  <- c(4, -1.91, 3.00, -5.002)
income <-  c(72.19, 1030.18, 10291.93, 1189192.18)
p_values <- c(0.12, 0.98, 0.0000191, 0.00000000002)

# Format c(0.0011, 0.011, 1) with digits = 1
format(c(0.0011, 0.011, 1),  digits = 1)

# Format c(1.0011, 2.011, 1) with digits = 1
format(c(1.0011, 2.011, 1),  digits = 1)

# Format percent_change to one place after the decimal point
format(percent_change, digit = 2)

# Format income to whole numbers
format(income, digits = 2)

# Format p_values in fixed format
format(p_values, scientific = F)
```

Not only does format() control the way the number is represented, it also controls some of the properties of the resulting string that affect its display.

```{r}
# Some vectors of numbers
percent_change  <- c(4, -1.91, 3.00, -5.002)
income <-  c(72.19, 1030.18, 10291.93, 1189192.18)
p_values <- c(0.12, 0.98, 0.0000191, 0.00000000002)

# Format c(0.0011, 0.011, 1) with digits = 1
format(c(0.0011, 0.011, 1),  digits = 1)

# Format c(1.0011, 2.011, 1) with digits = 1
format(c(1.0011, 2.011, 1),  digits = 1)

# Format percent_change to one place after the decimal point
format(percent_change, digit = 2)

# Format income to whole numbers
format(income, digits = 2)
```

The function formatC() provides an alternative way to format numbers based on C style syntax.

Rather than a scientific argument, formatC() has a format argument that takes a code representing the required format. The most useful are:

"f" for fixed,
"e" for scientific, and
"g" for fixed unless scientific saves space

```{r}
# From the format() exercise
x <- c(0.0011, 0.011, 1)
y <- c(1.0011, 2.011, 1)

# formatC() on x with format = "f", digits = 1
formatC(x, format = "f", digits = 1)

# formatC() on y with format = "f", digits = 1
formatC(y, format = "f", digits = 1)

# Format percent_change to one place after the decimal point
formatC(percent_change, format = "f", digits = 1)

# percent_change with flag = "+"
#The flag argument allows you to provide some modifiers that, for example, force
#the display of the sign (flag = "+"), left align numbers (flag = "-") and pad
#numbers with leading zeros (flag = "0"). You'll see an example in this exercise.

formatC(percent_change, format = "f", digits = 1, flag = "+")

# Format p_values using format = "g" and digits = 2
formatC(p_values, format = "g", digits = 2)
```

###### Putting strings together with paste()
```{r}
pretty_income <- c("72.19", "1030.18", "10291.93", "1189192.18")
pretty_percent <- c("+4.0", "-1.9", "+3.0", "-5.0")
# Add $ to pretty_income
paste("$", pretty_income, sep = "")
 
# Add % to pretty_percent
paste(pretty_percent, "%", sep = "")

# Create vector with elements like 2010: +4.0%`
years <- c(2010, 2011, 2012, 2013)
year_percent <- paste(years, sep = ": ", paste(pretty_percent,"%", sep = ""))

# Collapse all years into single string
paste(year_percent, collapse = ",")
```

###### A very simple table
```{r}
# Define the names vector
income <- c(72.19, 1030.18, 10291.93, 1189192.18)
income_names <- c("Year 0", "Year 1", "Year 2", "Project Lifetime")


# Create pretty_income by using format() with digits = 2 and big.mark = "," (income is loaded)
pretty_income <- format(income, digits = 2 , big.mark = ",")

# Create dollar_income
dollar_income <- paste("$", pretty_income, sep = "")

# Create formatted_names
formatted_names <- format(income_names, justify = "right")

# Create rows
rows <- paste(formatted_names, dollar_income, sep = "   ")

# Write rows
writeLines(rows)
```

###### another example
```{r}
toppings <- c("anchovies", "artichoke", "bacon", "breakfast bacon",
 "Canadian bacon", "cheese", "chicken", "chili peppers",
 "feta", "garlic", "green peppers", "grilled onions",  
"ground beef", "ham", "hot sauce", "meatballs", "mushrooms", "olives", "pineapple", "sausage", "spinach", "sun-dried tomato", "tomatoes")

# Randomly sample 3 toppings
my_toppings <- sample(toppings, size = 3)

# Print my_toppings
my_toppings

# Paste "and " to last element: my_toppings_and
my_toppings_and <- paste(c("", "", "and "), my_toppings, sep = "")

# Collapse with comma space: these_toppings
these_toppings <- paste(my_toppings_and, collapse = ", ")

# Add rest of sentence: my_order
my_order <- paste("I want to order a pizza with ", these_toppings, ".", sep = "")

# Order pizza with writeLines()
writeLines(my_order)
```



# Introducing stringr
__str_c()__ joins multiple strings into A single string.
```{r}
library(stringr)
#Let's look at str_c(), the c is short for concatenate, a function that works like paste()
my_toppings <- c("cheese", NA, NA)
my_toppings_and <- paste(c("", "", "and "), my_toppings, sep = "")
my_toppings_and

# Use str_c() instead of paste(): my_toppings_str
my_toppings_str <- str_c(c("", "", "and "), my_toppings)

# Print my_toppings_str
my_toppings_str
# Another stringr function that is useful when you may have missing values, is str_replace_na() which replaces missing values with any string you choose.

# paste() my_toppings_and with collapse = ", "
paste(my_toppings_and, collapse = ", ")
```

__str_length()__ takes a vector of strings as input and returns the number of characters in each string. 
```{r}

library(stringr)
library(babynames)
library(dplyr)

# Extracting vectors for boys' and girls' names
babynames_2014 <- filter(babynames, year == 2014)
boy_names <- filter(babynames_2014, sex == "M")$name
girl_names <- filter(babynames_2014, sex == "F")$name

# Take a look at a few boy_names
head(boy_names)

# Find the length of all boy_names
boy_length <- str_length(boy_names)

# Take a look at a few lengths
head(boy_length)

# Find the length of all girl_names
girl_length <- str_length(girl_names)

# Find the difference in mean length
mean(girl_length) - mean(boy_length)

# Confirm str_length() works with factors
head(str_length(factor(boy_names)))
```

__str_sub()__ extracts substrings from a character vector.
```{r}
#str_sub() function in stringr extracts parts of strings based on their location.
# Extract first letter from boy_names
boy_first_letter <- str_sub(boy_names, 1, 1)

# Tabulate occurrences of boy_first_letter
table(boy_first_letter)
  
# Extract the last letter in boy_names, then tabulate
boy_last_letter <- str_sub(boy_names, -1, -1)
table(boy_last_letter)

# Extract the first letter in girl_names, then tabulate
girl_first_letter <- str_sub(girl_names, 1, 1)
table(girl_first_letter)

# Extract the last letter in girl_names, then tabulate
girl_last_letter <- str_sub(girl_names, -1, -1)
table(girl_last_letter)
```

__str_detect()__ is used to answer the question: Does the string contain the pattern? It returns a logical vector of the same length as that of the input vector string. the result is alwyas TRUE or FALSE

```{r}
# Look for pattern "zz" in boy_names
contains_zz <- str_detect(boy_names, fixed("zz"))

# Examine str() of contains_zz
str(contains_zz)

# How many names contain "zz"?
sum(contains_zz)

# Which names contain "zz"?
boy_names[contains_zz]
```

__str_subset()__  gets a new vector of strings, but it only contains those original strings that contained the pattern. the results is a vector with all matching patterns
```{r}
#Find boy_names that contain "zz"
str_subset(boy_names, "zz")

# Find girl_names that contain "zz"
str_subset(girl_names, "zz")

# Find girl_names that contain "U"
starts_U <- str_subset(girl_names, "U")
starts_U

# Find girl_names that contain "U" and "z"
str_subset(starts_U, "z")
```

__str_count()__ answers the question "How many times does the pattern occur in each string?".
```{r}
# Count occurrences of "a" in girl_names
number_as <- str_count(girl_names, "a")

# Count occurrences of "A" in girl_names
number_As <- str_count(girl_names, "A")

# Histograms of number_as and number_As
hist(number_as)
hist(number_As)  

# Find total "a" + "A"
total_as <- number_as + number_As

# girl_names with more than 4 a's
girl_names[total_as > 4]
```

__str_split__ split the string into pieces and returns a list or  matrix if `simplify = TRUE`
```{r}
# Some date data
date_ranges <- c("23.01.2017 - 29.01.2017", "30.01.2017 - 06.02.2017")

# Split dates using " - ": CREATE A LIST
split_dates <- str_split(date_ranges, pattern = fixed(" - "))
split_dates
```

```{r}
# Some date data
date_ranges <- c("23.01.2017 - 29.01.2017", "30.01.2017 - 06.02.2017")

# Split dates using " - " CREATE A MATRIX
split_dates_n <- str_split(date_ranges, pattern = fixed(" - "), simplify = T, n = 2)
split_dates_n
```


```{r}
# From previous step
date_ranges <- c("23.01.2017 - 29.01.2017", "30.01.2017 - 06.02.2017")
split_dates_n <- str_split(date_ranges, fixed(" - "), n = 2, simplify = TRUE)

# Subset split_dates_n into start_dates and end_dates
start_dates <- split_dates_n[,1]

# Split start_dates into day, month and year pieces
str_split(start_dates, fixed("."),  simplify = TRUE)
```

```{r}
# From previous step
date_ranges <- c("23.01.2017 - 29.01.2017", "30.01.2017 - 06.02.2017")
split_dates_n <- str_split(date_ranges, fixed(" - "), n = 2, simplify = TRUE)

# Subset split_dates_n into start_dates and end_dates
start_dates <- split_dates_n[,1]

# Split start_dates into day, month and year pieces
str_split(start_dates, fixed("."),  simplify = TRUE)
```

```{r}
both_names <- c("Box, George", "Cox, David")

# Split both_names into first_names and last_names
both_names_split <- str_split(both_names, pattern = fixed(", "), simplify = TRUE)

# Get first names
first_names <- both_names_split[,2]
first_names

# Get last names
last_names <- both_names_split[,1]
last_names
```

Split lines into words where punctuations is included in the word
```{r}
# Split lines into words
words <- str_split(lines, pattern = c(" "))

# Number of words per line
lapply(words, length)
  
# Number of characters in each word
word_lengths <- lapply(words, str_length)
  
# Average word length per line
lapply(word_lengths,mean)

#The length() function returns the length of a vector (e.g. the number of words in each line) 
#the str_length() function returns the number of characters (e.g. in each word.)
```

Split lines into words where punctuations is NOT included in the word: __str_replace()__ ad __str_replace_all()__
```{r}
#ex 1: Some IDs
ids <- c("ID#: 192", "ID#: 118", "ID#: 001")

# Replace "ID#: " with ""
id_nums <- str_replace_all(ids, "ID#: ", "")

# Turn id_nums into numbers
id_ints <- as.numeric(id_nums)

#Ex 2: Some (fake) phone numbers
phone_numbers <- c("510-555-0123", "541-555-0167")

# Use str_replace() to replace "-" with " "
str_replace(phone_numbers, "-", " ")

# Use str_replace_all() to replace "-" with " "
str_replace_all(phone_numbers, "-", " ")

# Turn phone numbers into the format xxx.xxx.xxxx
str_replace_all(phone_numbers, "-", ".")
```

recap: 
```{r}
# Define some full namessubset()
names <- c("Diana Prince", "Clark Kent")

# Split into first and last names
names_split <- str_split(names, pattern = fixed(" "), simplify = TRUE)

# Extract the first letter in the first name
abb_first <- str_sub(names_split[,1], 1, 1)

# Combine the first letter ". " and last name
str_c(abb_first, ". ", names_split[,2])
```


# Pattern matching with regular expressions




a language for describing patterns
^.[\d]+
^ the start of the string, 
. followed by any single character, 
[\d]+ followed by one or more digits

START = 
END   = $
ANY_CHAR = .
DOT = \.

__str_view __ View HTML rendering of regular expression match.

```{r}
library(rebus.base)
# Some strings to practice with
x <- c("cat", "coat", "scotland", "tic toc")

# Match the strings that start with "co" 
str_view(x, pattern = START %R% "co")

# Match the strings that end with "at"
str_view(x, pattern = "at" %R% END)

# Match the string that is exactly "cat"
str_view(x, pattern = exactly("cat"))

# Match two characters, where the second is a "t"
str_view(x, pattern = ANY_CHAR %R% "t")

# Match a "t" followed by any character
str_view(x, pattern = "t" %R% ANY_CHAR)

# Match two characters
#Notice the first two characters are matched. Regular expression are lazy and will take the first match they find.
str_view(x, pattern = ANY_CHAR %R% ANY_CHAR)

# Match a string with exactly three characters
str_view(x, pattern = START %R% ANY_CHAR %R% ANY_CHAR %R% ANY_CHAR %R% END)

#Combine  regular expression with stringr
pattern <- "q" %R% ANY_CHAR

# Find names that have the pattern
names_with_q <- str_subset(boy_names, pattern = pattern)

# How many names were there?
length(names_with_q)

## Find part of name that matches pattern
part_with_q <- str_extract(boy_names, pattern)

# Get a table of counts
table(part_with_q)

## Did any names have the pattern more than once?
count_of_q <- str_count(boy_names, pattern)

# Get a table of counts
table(count_of_q)

##Get a logical vector of whether or not each boy's name contains q*
with_q <- str_detect(boy_names, pattern = "q" %R% ANY_CHAR)

# What fraction of babies got these names?
mean(with_q)

```

Alternation = (cat|dog)

pattern = char_class ("Aa") atch both __a__ pple and __A__ aron

pattern = negated_char_class ("Aa") match both a __p__ ple and Aa __r__ on

optional() = ?

Zero_or_more() = * 
one_or_more() = +
repeated() = {n}{m} - between n and M times

pattern = one_or_more("Aa") => __a__ pple, __Aa__ ron

```{r}
# Match Jeffrey or Geoffrey
whole_names <- or("Jeffrey","Geoffrey") 
str_view(boy_names, pattern = whole_names, match = TRUE)
#OR
# Match Jeffrey or Geoffrey, another way
common_ending <- or("Je", "Geo") %R% "ffrey"
str_view(boy_names, pattern = common_ending, match = TRUE)


# Match with alternate endings
by_parts <- or("Je", "Geo") %R% "ff" %R% or("ry", "ery", "rey", "erey")
str_view(boy_names, pattern = by_parts, match = TRUE)

# Match names that start with Cath or Kath
ckath <- or("C", "K") %R% "ath"
str_view(girl_names, pattern = ckath, match = TRUE)


#Create a character class that contains vowels, a, e, i, o, u and their upper case versions.
vowels <- char_class("Aa", "Ee", "Ii", "Oo", "Uu")

# Print vowels
vowels

# See vowels in x with str_view()
#only the first vowel is matched.
str_view(x, pattern = vowels)

# See vowels in x with str_view_all()
str_view_all(x, vowels)

```



Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.